---
title: "Realized Catch Model"
output:
  word_document: default
  html_notebook: default
---


## Introduction
A component of the operating model for the New England groundfish management strategy evaluation is a realized catch model.  This model applies an implementation error to convert the allowable catch limit for a stock into the realized catch of that stock.

In the case of a multispecies fisheries like the New England groundfish fisheries the catch of a stock may be a function not just of the ACL for that stock, but the catch or ACL for other stocks in the multispecies fishery.  We fit multivariate regressions to the available catch and ACL data as a means of predicting this interaction between stocks in the NE groundfish fishery.

## Data
Catch history data was downloaded from the Greater Atlantic Regional Fisheries Office year end catch accounting website for the New England multispecies (groundfish) fishery (https://www.greateratlantic.fisheries.noaa.gov/aps/monitoring/nemultispecies.html).  We downloaded data from 8 years of data available, for the 2010-2017 fiscal years, in the Year-End Groundfish Catch Accounting pdf files.

Data was extracted from these files and converted into a tidy usable format using R and the script in the wrangleCatchACL.R file.  This file extracts the five relevant data tables from each pdf, creating 5 data tables in R, one each for catch, ACL, the ratio of catch to ACL, landings, and discards, for 25 stocks:

```{r, echo=FALSE, message=FALSE}
library(tidyverse)
library(VGAM)
library(knitr)

##### - Functions - #####
# Spread keeping multiple value columns
myspread <- function(df, key, value) {
  # quote key
  keyq <- rlang::enquo(key)
  # break value vector into quotes
  valueq <- rlang::enquo(value)
  s <- rlang::quos(!!valueq)
  df %>% gather(variable, value, !!!s) %>%
    unite(temp, !!keyq, variable) %>%
    spread(temp, value)
}

catchHist<-read_csv("C:/Users/jcummings/OneDrive - UMASS Dartmouth/SMAST/Research/groundfish-MSE/data/data_processed/catchHistory/catchHist.csv")
catchHist <- catchHist %>% 
  select(Stock, Total, Year, data_type) %>% #select columns of interest
  rename_all(tolower) %>% #make all names lowercase
  mutate(stock = tolower(stock))
stocks<-data.frame(stocks=unique(catchHist$stock))
kable(stocks,caption="New England Multispecies stocks")
```

This results in the following data for each stock:
```{r, echo=FALSE, message=FALSE}
gbcod_data <- catchHist %>% 
  filter(grepl("gb cod",stock)) %>% 
  spread(data_type,total)
kable(gbcod_data,caption="George's Bank Cod data")
```


Plots of the George's Bank data, and subsets of the data showing the catch, ACL, and catch to ACL relationship through time:
```{r, echo=FALSE, message=FALSE}
mydata <- catchHist %>% 
  filter(grepl(paste(c("cod","haddock"), collapse="|"),stock))
ggplot(mydata, aes(x=year, y=total, group=data_type, color=data_type)) +
  geom_line() +
  facet_wrap(~stock,scales = "free") + labs(title="Time Series by stock and data type")

##### - Set up data to model cod-haddock Interactions - #####
# select just GB cod and haddock
mydata<-mydata %>% 
  filter(stock=="gb cod"|stock=="gb haddock")
mydata$stock[mydata$stock=="gb cod"]<-"cod"
mydata$stock[mydata$stock=="gb haddock"]<-"haddock"

# Plot time series of catch, ACL, C:ACL, Discards, and Landing
plotdata<-mydata %>% 
  filter(data_type=="Catch"|data_type=="ACL")

ggplot(plotdata, aes(x=year, y=total, group=data_type, color=data_type)) +
  geom_line() +
  facet_wrap(~stock,scales = "free") + labs(title="George's Bank Cod and Haddock ACL and Catch")

# Plot time series of catch to ACL for GB
plotdata2<-mydata %>% 
  filter(data_type=="CtoACL")

ggplot(plotdata2, aes(x=year, y=total, group=stock, color=stock)) +
  geom_line() + geom_hline(yintercept=100)+ ylab("Catch as Percentage of ACL") + labs(title="George's Bank Cod and Haddock Catch:ACL")

# organize for regression
regdata <- mydata %>%
  spread(data_type,total) %>% 
  mutate(ACL2=ACL^2) %>% 
  gather(data_type,total,-c(year,stock)) %>% 
  spread(stock,total) %>% 
  myspread(data_type, c("cod","haddock"))

```


## Realized Catch Modeling Methodology

Multiple regression models were fit to the George's Bank cod and George's Bank haddock catch, using George's Bank cod and haddock ACL as predictive variables.  We evaluated a set of five multiple regression models.  An intercept only model, a GB cod ACL model, a GB haddock ACL model, a GB cod + haddock ACL model, and a quadratic GB cod + haddock ACL model:
```{r, echo=TRUE, message=FALSE}
# Intercept only model
mvreg_int<-vglm(cbind(Catch_cod,Catch_haddock)~1,family=binormal,regdata)
# Cod ACL model
mvreg_cod<-vglm(cbind(Catch_cod,Catch_haddock)~ACL_cod,family=binormal,regdata)
# Haddock ACL model
mvreg_had<-vglm(cbind(Catch_cod,Catch_haddock)~ACL_haddock,family=binormal,regdata)
# Cod & Haddock model
mvreg_both<-vglm(cbind(Catch_cod,Catch_haddock)~ACL_cod+ACL_haddock,family=binormal,regdata)
# Quadratic Cod, & Haddock model
mvreg_both2<-vglm(cbind(Catch_cod,Catch_haddock)~ACL_cod+ACL2_cod+ACL_haddock+ACL2_haddock, family=binormal,regdata)
```

As a quick model selection procedure these five models were evaluated using AIC and AICc scores.  
```{r, echo=FALSE, message=FALSE}
# Calculate AIC scores
mvregAIC<-data.frame(model=c("Intercept","Cod ACL","Haddock ACL","Cod ACL & Haddock ACL","Quadratic Cod & Haddock ACL"),AIC=c(AIC(mvreg_int),AIC(mvreg_cod),AIC(mvreg_had),AIC(mvreg_both),AIC(mvreg_both2)),AICc=c(AICc(mvreg_int),AICc(mvreg_cod),AICc(mvreg_had),AICc(mvreg_both),AICc(mvreg_both2)))
kable(mvregAIC,caption="AIC results")
```

Based on the results of this quick analysis we progressed with the Cod ACL and Cod ACL & Haddock ACL models.  

###### Cod Model
This resulting model coefficients and predicted mean catch given a cod ACL are:
```{r, echo=FALSE, message=FALSE}
# coefficients
kable(Coef(mvreg_cod, matrix = TRUE),caption="Cod ACL model coefficients")

# predictions
##### - set up prediction for Cod ACL model- #####
cod_ACL<-seq(0,10000,length.out=21)
cod_pred<-predictvglm(mvreg_cod,list(ACL_cod=cod_ACL)) %>% 
  as_tibble() %>% 
  select(mean1,mean2) %>% 
  mutate(cod=mean1) %>% 
  mutate(haddock=mean2) %>%
  select(cod,haddock)
cod_pred<-cbind(cod_ACL,cod_pred)
kable(cod_pred,caption="Cod model mean predicted catch")
```

The mean predicted relationship between the cod ACL and the cod and haddock catch from the Cod ACL model is:
```{r, echo=FALSE, message=FALSE}
# plot
pred_plot<-cod_pred %>%
  gather(catch,value, -cod_ACL)

# plot prediction
ggplot(pred_plot, aes(x=cod_ACL,y=value,group=catch,col=catch)) +
  geom_line() + xlab("Cod ACL") + ylab("Catch")
```

###### Cod & Haddock Model
This resulting model coefficients and predicted mean catch given a cod ACL and haddock ACL are:
```{r, echo=FALSE, message=FALSE}
kable(Coef(mvreg_both, matrix = TRUE),,caption="Cod & Haddock ACL model coefficients")

haddock_ACL<-seq(0,100000,length.out=21)
codhad_pred<-predictvglm(mvreg_both,list(ACL_cod=cod_ACL,ACL_haddock=haddock_ACL)) %>% 
  as_tibble() %>% 
  select(mean1,mean2) %>% 
  mutate(cod=mean1) %>% 
  mutate(haddock=mean2) %>%
  select(cod,haddock)
codhad_pred<-cbind(cod_ACL,haddock_ACL,codhad_pred)
kable(codhad_pred,caption="Cod model mean predicted catch")
```

```{r, echo=FALSE, message=FALSE}
##### - Set up prediction for Cod and Haddock ACL model - #####
# Create ACL combos
pred_vals<-expand.grid(cod_ACL,haddock_ACL)
# Predict
pred_plot_both<-predictvglm(mvreg_both,list(ACL_cod=pred_vals$Var1,ACL_haddock=pred_vals$Var2))
# reorganize data
pred_plot_both2<-pred_plot_both %>%
  as_tibble() %>% 
  select(mean1,mean2) %>% 
  mutate(Catch_Cod=mean1) %>% 
  mutate(Catch_Haddock=mean2) %>%
  select(Catch_Cod,Catch_Haddock) %>% 
  cbind(Cod_ACL=pred_vals$Var1,Haddock_ACL=pred_vals$Var2) %>% 
  gather(catch,value, -Cod_ACL, -Haddock_ACL)
```

The mean predicted relationship between the cod ACL and the cod and haddock catch from the Cod ACL model is:
```{r, echo=FALSE, message=FALSE}
# Plot
ggplot(pred_plot_both2, aes(x=Cod_ACL,y=Haddock_ACL,col=value)) + 
  facet_grid(~catch) +  geom_point(size=6)+geom_contour(aes(z=value),col="black") + theme_bw()
```
