---
title: 'Implementation Error'
author: Birkenbach, Truesdell, Cummings(t), Kerr, Lee, Fay(t)  
csl: canadian-journal-of-fisheries-and-aquatic-sciences.csl
output:
  pdf_document:
    includes:
      in_header: preamble-latex.tex
    keep_tex: yes
    pandoc_args: --latex-engine=xelatex
  html_document: null
  word_document: null
bibliography: minyang_bib.bib
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.align='center')
if(!require(ggplot2)) {  
  install.packages("ggplot2")
  require(ggplot2)}
if(!require(tidyr)) {  
  install.packages("tidyr")
  require(tidyr)}
if(!require(dplyr)) {  
  install.packages("dplyr")
  require(dplyr)}
if(!require(broom)) {  
  install.packages("broom")
  require(broom)}
if(!require(knitr)) {  
  install.packages("knitr")
  require(knitr)}
if(!require(gridExtra)) {  
  install.packages("gridExtra")
  require(gridExtra)}
if(!require(grid)) {  
  install.packages("grid")
  require(grid)}
if(!require(ggthemes)) {  
  install.packages("ggthemes")
  require(ggthemes)}


#to get table captions and numbering
tn = local({
  i = 0
  function(x) {
    i <<- i + 1
    #paste('\n\n:Table ', i, ': ', x, sep = '')
    paste('\n\n:', x, sep = '')
    # The : before Table tells pandoc to wrap your caption in <caption></caption>
  }
})
knit_hooks$set(tab.cap = function(before, options, envir) {
  if(!before)
    tn(options$tab.cap)
})
default_output_hook = knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  if (is.null(options$tab.cap) == F)  
    x
  else
    default_output_hook(x,options)
})

#a better plotting theme than ggplot default? from https://rpubs.com/Koundy/71792
theme_Publication <- function(base_size=14, base_family="") {
      library(grid)
      library(ggthemes)
      (theme_foundation(base_size=base_size, base_family=base_family)
       + theme(plot.title = element_text(face = "bold",
                                         size = rel(1.2), hjust = 0.5),
               text = element_text(),
               panel.background = element_rect(colour = NA),
               plot.background = element_rect(colour = NA),
               panel.border = element_rect(colour = NA),
               axis.title = element_text(face = "bold",size = rel(1)),
               axis.title.y = element_text(angle=90,vjust =2),
               axis.title.x = element_text(vjust = -0.2),
               axis.text = element_text(), 
               axis.line = element_line(colour="black"),
               axis.ticks = element_line(),
               panel.grid.major = element_line(colour="#f0f0f0"),
               panel.grid.minor = element_blank(),
               legend.key = element_rect(colour = NA),
               legend.position = "right",
               legend.direction = "vertical",
               legend.key.size= unit(0.4, "cm"),
               legend.margin = unit(0, "cm"),
               legend.title = element_text(face="italic"),
               plot.margin=unit(c(10,5,5,5),"mm"),
               strip.background=element_rect(colour="#f0f0f0",fill="#f0f0f0"),
               strip.text = element_text(face="bold")
          ))
      
}

scale_fill_Publication <- function(...){
      library(scales)
      discrete_scale("fill","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}

scale_colour_Publication <- function(...){
      library(scales)
      discrete_scale("colour","Publication",manual_pal(values = c("#386cb0","#fdb462","#7fc97f","#ef3b2c","#662506","#a6cee3","#fb9a99","#984ea3","#ffff33")), ...)

}
```

\doublespacing

<!---Oh god this is such a hack -->

A. Birkenbach: University of Delaware, College of Earth, Ocean, and Environment Street, 207 Robinson Hall, Newark , DE 19716. abirken@udel.edu tel: (302) 831-4664 Fax: (XXX) YYY-ZZZZ

M. Lee: NOAA Fisheries, Northeast Fisheries Science Center. 166 Water St. Woods Hole MA. 02543.  Min-Yang.Lee@noaa.gov 

O. Authors:  

This draft manuscript is distributed solely for purposes of scientific peer review.

Funding acknowledgements:
\newpage


#Abstract
To be reasonably accurate, a multispecies simulation model must account for techincal interactions. We do that for the northeast US groundfish fishery.  This allows for fishing mortality to be systematically different from the level set by fishery managers.

#Introduction
Many simulation models used to inform fisheries managers take a phenomenological (reduced-form) approach [@Fenichel2013] to the process by which humans kill fish.  For example, @punt2008evaluation, @Deroba2012Evaluating, and @Deroba2019 assume that the actual fishing mortality rate is equal to the fishing mortality rate set by the manager as modified by an idiosycratic error term.  The distribution of this error term is approximated from historical data.  @Fulton2011a note that implementation error encompassesto three distinct processes: (1) decisions where scientific advice on stock status is subordinated; (2) lack of direct regulatory control of fishing mortality due to inadequate enforcement or a lack of a direct link between controls (regulations) and harvest; and (3) unexpected behavioural responses of resource users.  

We aim to address (2) and (3); we will not address (1).  We pair a mechanistic (structural) model of production in a multispecies fishery with stock dynamics and assessment models. Our general approach is similar to that of @Ono2018.  We focus on one specific way that technical interactions (or jointness in production) matter: the bycatch or "choke stock" story[^choke]  in multispecies fishery.[^diffs] We apply the model to the northeast US multispecies fishery, commonly referred to as groundfish.  Participants in this fishery have been observed to change their  behavior in response to recent policy changes: @Scheld2018 and @Holzer2019 document changes in fishing behavior after changes in management.


@Ono2018 states:
>Some MSE frameworks include some elements of fleet dynamics (e.g., Atlantis (Fulton et al. 2011), BIOMAS (Ives et al. 2013), FLR (Kell et al. 2007), and ISIS-FISH (Pelletier et al. 2009)), but a multi- species fishery quota allocation model with trade-offs and con- straints (i.e., based on constrained optimization) as well as the temporal dynamics of technical interactions has never been included in an MSE.

[^choke]: we are not actually going to call this the "choke stock" story.

[^diffs]: We're doing something pretty different from @Ono2018.  Structural model instead of metier model. No "expansion factor" hacks for behavioral changes.  Fish stock model is fine.  

Single species models with random implementation error cannot credibly model interrelated fisheries with technical interactions. 

We realistically allow for leakages out of the main fishery into other fisheries [@Cunningham2016].  Because our model is grounded in a structural model of individuals's utility maximization, instead of a reduced form model of metiers, we are better able to predict changes in behavior that result from changes in policies.  For example, we can examine policies that include tradable output controls (ITQs), non-tradable output controls (IFQs), and certain input controls (Days-at-Sea).  We also do some other cool stuff.

What's the take-home? Even without an ecological linkage (predator-prey or competition), a single species model of $F$ with implementation error lead to very recommendations compared to a multispecies model?



<!---
The implementation error approach is quite common.

This is a common approach(Irwin et al 2008). 
(Rosenberg and Brault 1993 Rosenberg, A.A. and Brault, S. (1993) Choosing a man-
agement strategy for stock rebuilding when control is
uncertain. Canadian Special Publication of Fisheries and
Aquatic Sciences 120, 243–253.

O’Boyle 1993  O’Boyle, R. (1993) Fisheries management organizations: a
study in uncertainty. Canadian Special Publication of
Fisheries and Aquatic Sciences 120, 423–436.

FAO 1996  FAO (1996) Precautionary approach to fisheries Part 2:
scientific papers prepared for the technical consultation
on the precautionary approach to capture fisheries
(including species introductions) Lysekil, Sweden, 6–13
June 1995. FAO Fish Tech Paper 350(Part 2). FAO,
Rome. p. 210.

 Francis and Shotton rancis, R.I.C.C. and Shotton, R. (1997) ‘‘Risk’’ in fisheries
management: a review. Canadian Journal of Fisheries and
Aquatic Sciences 54, 1699–1715.

Dichemont 2006abc papers.
 -->

## Related Literature
<!---We're not actually going to have a related literature section -->
Construction of complex simulation models with many sources of uncertainty is standard in the fisheries literature [c.f. @Smith1999, @Punt2016].  Our model is similar to that of @Ono2018, which models three principal groundfish stocks (Pacific Cod, Alaska Pollock, and Yellowfin Sole) and one bycatch species (Pacific Halibut).  Fishing vessels are assigned to metiers and assigned a baseline catch level based on historical data.  The fishery manager *selects* total allowable catch for each metier-species combination. SSB will scale the baseline catch. It's unclear how/why the fishery stops fishing. But it's also quasi-behavioral, there's an "adjustment" to the scaling that accounts for adaptation.


#Methods
Here is the methods section

## A structural economic model of targeting 

## A model of the fishery 

## Performance Metrics

### Economic Performance Metrics

# Results
Here is an overview of our results.


# Discussion
Here is the discussion of our results





# Acknowledgements

# References
<div id="refs"></div>

\newpage

# Tables

<!---markdown style table. fucking escape characters-->
```{r symboltable, tab.cap="Table of Symbols \\label{symtab}", echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
tabl <- "
|	Symbol	|	Definition
|:--------	|:--------------------------------------------------------
|	$y$	|	year
|	$a$	|	age
|	$\\varepsilon_{\\phi,y}$	|	assessment error
|	$\\bar{F}_y$	|	fishing mortality that would remove quota from herring population
|	$Q_y$	|	quota for herring fishery
|	$\\varepsilon_{\\theta,y}$	|	implementation error
|	$\\sigma_{\\theta}^2$	|	variance of implementation error
|	$GR_y$	|	gross revenue
|	$q_y$	|	quantity landed
"
cat(tabl) # output the table in a format good for HTML/PDF/docx conversion
```


# Figures




